---
title: "Short-term impacts of island restoration on seabirds in Gwaii Haanas"
knit: (function(inputFile, encoding) { 
      out_dir <- 'reports';
      ext <- '.html';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir,  paste('GH-acoustic-Report-', lubridate::date(Sys.time()), ext, sep = ''))) })
output:
  html_document:
    fig_caption: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F,message=F,results="hide", prompt=F, error=F)
```

```{r}
source('header.R')
source('voc-analysis.R')
source('voc-plots.R')
source('voc-captions.R')
source('spatial.R')
```

## Spatial and temporal coverage
Task 2a) `r task2a`


`r figs("spatnorth")`
```{r spatnorth, fig_height = 5}
mapnorth
```

`r figs("spatsouth")`
```{r spatsouth, fig_height = 4}
mapsouth
```

`r figs("tempcov")`
```{r tempcov, fig_height = 4}
tempcov
```

# Vocalization Probability
Task 2c) `r task2c`

## Methods
Acoustic Recording Units (ARUs) were located at sites to target breeding habitat of certain species. Some sites had no or very few detections of certain species. For description of vocalization probabilities for each species, we included only those sites deemed to have substantial vocalization activity, defined as any site with a proportion of recordings with detection > 0.01, for a particular species. This approach was taken rather than relying on a priori understanding of habitat and colony locations. For Ancient Murrelets, Cassin's Auklets, Fork-Tailed Storm-Petrels, and Leach's Storm-Petrels: `r ddply(sitedet, .(species), summarise, length = length(species))$length[1]`, `r ddply(sitedet, .(species), summarise, length = length(species))$length[2]`, `r ddply(sitedet, .(species), summarise, length = length(species))$length[3]` and `r ddply(sitedet, .(species), summarise, length = length(species))$length[4]` sites had 50 or more detections, respectively. To describe hourly vocalisation probabilities, we calculated: (1) proportion of recordings with detection of at least one individual in each 10-minute recording period ('p(presence)'); and (2) proportion of recordings with detection of at least two individuals in each recording period ('p(presence > 1)'). We calculated the same metrics (presence, presence of multiple individuals) aggregated by night, for description of nighly vocalization probability. For boxplots, voc.night.prop1proportions were calculated seprately for each site included, which allows visualization fo the variation between sites. We also provide plots that show the mean of proportions at all sites included, with binomial 95% confidence intervals. For description of nightly vocalization probability, separate plots were provided for all years combined, and for each year individually to visually assess whether patterns in vocalization probability varied between years.

## Results

### Daily vocalization probabilities

`r figs("voc.night.m")`

```{r voc.night.m}
point(voc.night.m, 'hrmin', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M')
```

`r figs("voc.night.cov.site")`

```{r, voc.night.cov.site, fig.height = 2 "Number of sites in each 10-minute recording period.")}
coverage(voc.night.cov, 'hrmin', 'sites', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M')
```

### Seasonal vocalization probabilities, all years combined

`r figs("voc.allyear.m")`

```{r voc.allyear.m}
point(voc.allyear.m, 'ynight', 'species')
```

`r figs("voc.allyear.cov.site")`

```{r, voc.allyear.cov.site, fig.height = 2, fig.width = 7}
coverage(voc.allyear.cov, 'ynight', 'sites', 'species')

```

##### Difference between responses

`r figs("voc.allyear.m")`

```{r voc.allyear.m}
ggplot(voc.allyear.diff) +
    geom_point( aes_string(x = 'ynight', y = 'meandiff'), size = 0.4)  + 
    geom_line(aes_string(x = 'ynight', y = 'meandiff'), size = 0.3) +
    facet_wrap(~species, ncol = 2, strip.position = 'top') + theme_bw() +
    scale_x_datetime(breaks = date_breaks('1 month'), labels = date_format('%b')) + 
    labs(y = 'Response difference', x = 'Date')  
```


### Seasonal vocalization probabilities by year
#### Ancient Murrelet

`r figs("voc.byyear.m.anmu")`

```{r voc.byyear.m.anmu}
point(voc.byyear.anmu.m, 'ynight', 'year')
```

`r figs("voc.byyear.cov.anmu.site")`

```{r, voc.byyear.cov.anmu.site, fig.height = 3.5}
coverage(voc.byyear.anmu.cov, 'ynight', 'sites', 'species')
```

#### Cassin's Auklet
`r figs("voc.byyear.m.caau")`

```{r voc.byyear.m.caau}
point(voc.byyear.caau.m, 'ynight', 'year')
```

`r figs("voc.byyear.cov.caau.site")`

```{r, voc.byyear.cov.caau.site, fig.height = 3.5}
coverage(voc.byyear.caau.cov, 'ynight', 'sites', 'species')
```

#### Fork-Tailed Storm-Petrel
`r figs("voc.byyear.m.ftsp")`

```{r voc.byyear.m.ftsp}
point(voc.byyear.ftsp.m, 'ynight', 'year')
```

`r figs("voc.byyear.cov.ftsp.site")`

```{r, voc.byyear.cov.ftsp.site, fig.height = 3.5}
coverage(voc.byyear.ftsp.cov, 'ynight', 'sites', 'species')
```


#### Leach's Storm-Petrel
`r figs("voc.byyear.m.lesp")`

```{r voc.byyear.m.lesp}
point(voc.byyear.lesp.m, 'ynight', 'year')
```

`r figs("voc.byyear.cov.lesp.site")`

```{r, voc.byyear.cov.lesp.site, fig.height = 3.5}
coverage(voc.byyear.lesp.cov, 'ynight', 'sites', 'species')
```

#### Seasonal Vocalisation - 14-day sliding window means

`r figs("voc.allyear.m14")`

```{r, voc.allyear.m14, fig.height = 3.5}
window14
```

#### Nightly Vocalisation - 1 hour sliding means

`r figs("voc.night.m3")`

```{r, voc.night.m3, fig.height = 3.5}
window3
```

## Appendix A - Coverage by recordings
`r figs("voc.night.cov.rec")`

```{r, voc.night.cov.rec, fig.height = 2}
coverage(voc.night.cov, 'hrmin', 'recordings', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M')
```

`r figs("voc.allyear.cov.rec")`

```{r, voc.allyear.cov.rec, fig.height = 2}
coverage(voc.allyear.cov, 'ynight', 'recordings', 'species', 'Total recordings', 'p(presence)', '1 month', '%b')
```

`r figs("voc.byyear.cov.anmu.rec")`

```{r, voc.byyear.cov.anmu.rec}
coverage(voc.byyear.anmu.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b')
```

`r figs("voc.byyear.cov.caau.rec")`

```{r, voc.byyear.cov.caau.rec}
coverage(voc.byyear.caau.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b')
```

`r figs("voc.byyear.cov.ftsp.rec")`

```{r, voc.byyear.cov.ftsp.rec}
coverage(voc.byyear.ftsp.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b')
```

`r figs("voc.byyear.cov.lesp.rec")`

```{r, voc.byyear.cov.lesp.rec}
coverage(voc.byyear.lesp.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b')
```

<!-- ```{r voc.night, fig.cap = fig$cap("voc.night", "Daily vocalization probabilities by species. Each boxplot summarizes proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot.n(voc.night, 'hrmin', 'prop1', 'hrmin', 'species', 'p(presence)') -->
<!-- ``` -->

<!-- ```{r voc.night.prop2, fig.cap = fig$cap("voc.night.prop2", "Daily vocalization probabilities by species. Each boxplot summarizes proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot.n(voc.night, 'hrmin', 'prop2', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M') -->
<!-- ``` -->

<!-- ```{r, voc.allyear.prop1, fig.cap = fig$cap("voc.allyear.prop1", "Seasonal vocalization probabilities by species, all years combined. Each boxplot summarizes proportion of recordings with detection of one or more individuals at each site.")} -->
<!-- boxplot.y(voc.allyear.w, 'week', 'prop1', 'week', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.allyear.prop2, fig.cap = fig$cap("voc.allyear.prop2", "Seasonal vocalization probabilities by species, all years combined. Each boxplot summarizes proportion of recordings with detection of two or more individuals at each site.")} -->
<!-- boxplot(voc.allyear.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.anmu, fig.cap = fig$cap("voc.byyear.anmu", "Seasonal vocalization probabilities of Ancient Murrelet, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.anmu.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.anmu.prop2, fig.cap = fig$cap("voc.byyear.anmu.prop2", "Seasonal vocalization probabilities of Ancient Murrelet, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.anmu.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.caau, fig.cap = fig$cap("voc.byyear.caau", "Seasonal vocalization probabilities of Cassin's Auklet, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.caau.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.caau.prop2, fig.cap = fig$cap("voc.byyear.caau.prop2", "Seasonal vocalization probabilities of Cassin's Auklet, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.caau.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.ftsp, fig.cap = fig$cap("voc.byyear.ftsp", "Seasonal vocalization probabilities of Fork-Tailed Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.ftsp.prop2, fig.cap = fig$cap("voc.byyear.ftsp.prop2", "Seasonal vocalization probabilities of Fork-Tailed Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.lesp, fig.cap = fig$cap("voc.byyear.lesp", "Seasonal vocalization probabilities of Leach's Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.lesp.prop2, fig.cap = fig$cap("voc.byyear.lesp.prop2", "Seasonal vocalization probabilities of Leach's Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

