---
title: "Short-term Impacts of Island Restoration on Seabirds in Gwaii Haanas: Coverage and Vocalisation Probability"
knit: (function(inputFile, encoding) { 
      out_dir <- 'reports';
      ext <- '.html';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir,  paste('GH-acoustic-Report-Voc-', lubridate::date(Sys.time()), ext, sep = ''))) })
output:
  html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo=F, warning=F,message=F,results="hide", prompt=F, error=F, fig.path = 'reports/files/')
```

```{r}
source('header.R')
lapply(list.files(path = 'data', pattern = "*.Rda", full.names = TRUE), load, .GlobalEnv)
source('voc-plots.R')
source('voc-captions.R')
source('spatial.R')
```

## Spatial and temporal coverage
Task 2a) `r task2a`


`r figs("spatnorth.ph2")`

```{r spatnorth, fig_height = 5}
mapnorth_ph2()
```

`r figs("spatsouth.ph2")`

```{r spatsouth, fig_height = 4}
mapsouth_ph2()
```

`r figs("tempcov.ph2")`

```{r tempcov, fig.height = 5, fig.width = 10}
tempcov(data = filter(sea, phase2 == 1))
```

# Vocalization Probability
Task 2c) `r task2c`

## Methods
Acoustic Recording Units (ARUs) were located at sites to target breeding habitat of certain species. Some sites had no or very few detections of certain species. For description of vocalization probabilities by species, we included only those sites deemed to have high vocalization activity, defined as any site with a proportion of recordings with detection > 0.01, for a particular species. For Ancient Murrelets, Cassin's Auklets, Fork-Tailed Storm-Petrels, and Leach's Storm-Petrels: `r ddply(sitedet, .(species), summarise, length = length(species))$length[1]`, `r ddply(sitedet, .(species), summarise, length = length(species))$length[2]`, `r ddply(sitedet, .(species), summarise, length = length(species))$length[3]` and `r ddply(sitedet, .(species), summarise, length = length(species))$length[4]` sites had > 0.01 proportion of recordings with detection, respectively. To describe hourly vocalisation probabilities, we calculated: (1) proportion of recordings with detection of at least one individual in each 10-minute recording period ('p(presence)'); and (2) proportion of recordings with detection of at least two individuals in each recording period ('p(presence > 1)'). We calculated the same metrics (p(presence), p(presence>1)) aggregated by night, for description of seasonal vocalization probability. For visulaisation, the mean of all sites within each time period was calculated, along with standard error intervals. For boxplots, proportions were calculated for each week (Appendix A). For description of seasonal vocalization probability, separate plots were provided for all years combined, and for each year individually to visually assess whether patterns in vocalization probability varied between years. However, temporal coverage was limited in some years, which hindered comparisons.

## Results

### Daily vocalization probabilities

`r figs("voc.night.m")`

```{r voc.night.m}
point(voc.night.m, 'hrmin', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M')
```

```{r voc.night.m.pa}
point_pa(voc.night.m, 'hrmin', 'species', 'Time of Night', '1 hour', '%H:%M')
```

`r figs("voc.night.cov.site")`

```{r, voc.night.cov.site, fig.height = 2}
coverage(voc.night.cov, 'hrmin', 'sites', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M')
```

`r figs("window3")`

```{r, window3}
ggplot(filter(voc.night.m, response == 'p(presence)'), aes(label = paste(hour(hrmin), lubridate::minute(hrmin), sep = ':'))) +
  geom_line(aes(x = hrmin, y = mean.3, color = species)) + 
  geom_point(aes(x = hrmin, y = max.3), color = 'black') + 
  geom_label(aes(x = hrmin, y = max.3), hjust = -0.2, vjust = 0.2, size = 3) +  
  labs(x = 'Time of Night', y = 'p(presence)', color = 'Species') +
  scale_x_datetime(breaks = date_breaks('2 hours'), labels = date_format("%H:%M", tz = tz_data))  + theme_bw() + theme_big
```

### Seasonal vocalization probabilities, all years combined

`r figs("voc.allyear.m")`

```{r voc.allyear.m}
point(voc.allyear.m, 'ynight', 'species')
```

```{r voc.allyear.m.pa}
point_pa(voc.allyear.m, 'ynight', 'species')
```

`r figs("voc.allyear.cov.site")`

```{r, voc.allyear.cov.site, fig.height = 2, fig.width = 7}
coverage(voc.allyear.cov, 'ynight', 'sites', 'species')

```

`r figs("window14")`

```{r, window14}
ggplot(filter(voc.allyear.m, response == 'p(presence)'), aes(label = paste(month(ynight, label = TRUE), day(ynight), sep = '-'))) + 
  geom_line(aes(x = ynight, y = mean.14, color = species)) +
  geom_point(aes(x = ynight, y = max.14), color = 'black') + 
  geom_label(aes(x = ynight, y = max.14), hjust = -0.2, vjust = 0.2, size = 3) + theme_bw() +
  labs(x = 'Date', y = 'p(presence)', color = 'Species') +
  scale_x_datetime(breaks = date_breaks('1 month'), labels = date_format("%b")) + theme_big
```

`r figs("voc.allyear.diff")`

```{r voc.allyear.diff}
ggplot(voc.allyear.diff) +
    geom_point( aes_string(x = 'ynight', y = 'meandiff'), size = 0.4)  + 
    geom_line(aes_string(x = 'ynight', y = 'meandiff'), size = 0.3) +
    facet_wrap(~species, ncol = 2, strip.position = 'top') + theme_bw() +
    scale_x_datetime(breaks = date_breaks('1 month'), labels = date_format('%b')) + 
    labs(y = 'p(presence) - p(presence>1)', x = 'Date')   + theme_big
```


### Seasonal vocalization probabilities, by year
#### Ancient Murrelet

`r figs("voc.byyear.m.anmu")`

```{r voc.byyear.m.anmu}
point(filter(voc.byyear.m, species == sp[1]), 'ynight', 'year')
```

```{r voc.byyear.m.anmu.pa}
point_pa(filter(voc.byyear.m, species == sp[1]), 'ynight', 'year')
```

`r figs("voc.byyear.cov.anmu.site")`

```{r, voc.byyear.cov.anmu.site, fig.height = 3.5}
coverage(filter(voc.byyear.cov, species == sp[1]), 'ynight', 'sites', 'year', ncol = 3)
```

#### Cassin's Auklet

`r figs("voc.byyear.m.caau")`

```{r voc.byyear.m.caau}
point(filter(voc.byyear.m, species == sp[2]), 'ynight', 'year')
```

```{r voc.byyear.m.caau.pa}
point_pa(filter(voc.byyear.m, species == sp[2]), 'ynight', 'year')
```

`r figs("voc.byyear.cov.caau.site")`

```{r, voc.byyear.cov.caau.site, fig.height = 3.5}
coverage(filter(voc.byyear.cov, species == sp[2]), 'ynight', 'sites', 'year', ncol = 3)
```

#### Fork-Tailed Storm-Petrel

`r figs("voc.byyear.m.ftsp")`

```{r voc.byyear.m.ftsp}
point(filter(voc.byyear.m, species == sp[3]), 'ynight', 'year')
```

```{r voc.byyear.m.ftsp.pa}
point_pa(filter(voc.byyear.m, species == sp[3]), 'ynight', 'year')
```

`r figs("voc.byyear.cov.ftsp.site")`

```{r, voc.byyear.cov.ftsp.site, fig.height = 3.5}
coverage(filter(voc.byyear.cov, species == sp[3]), 'ynight', 'sites', 'year', ncol = 3)
```

#### Leach's Storm-Petrel

`r figs("voc.byyear.m.lesp")`

```{r voc.byyear.m.lesp}
point(filter(voc.byyear.m, species == sp[4]), 'ynight', 'year')
```

```{r voc.byyear.m.lesp.pa}
point_pa(filter(voc.byyear.m, species == sp[4]), 'ynight', 'year')
```

`r figs("voc.byyear.cov.lesp.site")`

```{r, voc.byyear.cov.lesp.site, fig.height = 3.5}
coverage(filter(voc.byyear.cov, species == sp[4] ), 'ynight', 'sites', 'year', ncol = 3)
```

<!-- ## Appendix A - Coverage by recordings -->
<!-- `r figs("voc.night.cov.rec")` -->

<!-- ```{r, voc.night.cov.rec, fig.height = 2} -->
<!-- coverage(voc.night.cov, 'hrmin', 'recordings', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M') -->
<!-- ``` -->

<!-- `r figs("voc.allyear.cov.rec")` -->

<!-- ```{r, voc.allyear.cov.rec, fig.height = 2} -->
<!-- coverage(voc.allyear.cov, 'ynight', 'recordings', 'species', 'Total recordings', 'p(presence)', '1 month', '%b') -->
<!-- ``` -->

<!-- `r figs("voc.byyear.cov.anmu.rec")` -->

<!-- ```{r, voc.byyear.cov.anmu.rec} -->
<!-- coverage(voc.byyear.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b') -->
<!-- ``` -->

<!-- `r figs("voc.byyear.cov.caau.rec")` -->

<!-- ```{r, voc.byyear.cov.caau.rec} -->
<!-- coverage(voc.byyear.caau.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b') -->
<!-- ``` -->

<!-- `r figs("voc.byyear.cov.ftsp.rec")` -->

<!-- ```{r, voc.byyear.cov.ftsp.rec} -->
<!-- coverage(voc.byyear.ftsp.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b') -->
<!-- ``` -->

<!-- `r figs("voc.byyear.cov.lesp.rec")` -->

<!-- ```{r, voc.byyear.cov.lesp.rec} -->
<!-- coverage(voc.byyear.lesp.cov, 'ynight', 'sites', 'species', 'Date', 'Total sites', '1 month', '%b') -->
<!-- ``` -->

<!-- ```{r voc.night, fig.cap = fig$cap("voc.night", "Daily vocalization probabilities by species. Each boxplot summarizes proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot.n(voc.night, 'hrmin', 'prop1', 'hrmin', 'species', 'p(presence)') -->
<!-- ``` -->

<!-- ```{r voc.night.prop2, fig.cap = fig$cap("voc.night.prop2", "Daily vocalization probabilities by species. Each boxplot summarizes proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot.n(voc.night, 'hrmin', 'prop2', 'species', 'Time of Night', 'p(presence)', '1 hour', '%H:%M') -->
<!-- ``` -->

<!-- ```{r, voc.allyear.prop1, fig.cap = fig$cap("voc.allyear.prop1", "Seasonal vocalization probabilities by species, all years combined. Each boxplot summarizes proportion of recordings with detection of one or more individuals at each site.")} -->
<!-- boxplot.y(voc.allyear.w, 'week', 'prop1', 'week', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.allyear.prop2, fig.cap = fig$cap("voc.allyear.prop2", "Seasonal vocalization probabilities by species, all years combined. Each boxplot summarizes proportion of recordings with detection of two or more individuals at each site.")} -->
<!-- boxplot(voc.allyear.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.anmu, fig.cap = fig$cap("voc.byyear.anmu", "Seasonal vocalization probabilities of Ancient Murrelet, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.anmu.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.anmu.prop2, fig.cap = fig$cap("voc.byyear.anmu.prop2", "Seasonal vocalization probabilities of Ancient Murrelet, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.anmu.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.caau, fig.cap = fig$cap("voc.byyear.caau", "Seasonal vocalization probabilities of Cassin's Auklet, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.caau.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.caau.prop2, fig.cap = fig$cap("voc.byyear.caau.prop2", "Seasonal vocalization probabilities of Cassin's Auklet, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.caau.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.ftsp, fig.cap = fig$cap("voc.byyear.ftsp", "Seasonal vocalization probabilities of Fork-Tailed Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.ftsp.prop2, fig.cap = fig$cap("voc.byyear.ftsp.prop2", "Seasonal vocalization probabilities of Fork-Tailed Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.lesp, fig.cap = fig$cap("voc.byyear.lesp", "Seasonal vocalization probabilities of Leach's Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of one or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop1', 'species') -->
<!-- ``` -->

<!-- ```{r, voc.byyear.lesp.prop2, fig.cap = fig$cap("voc.byyear.lesp.prop2", "Seasonal vocalization probabilities of Leach's Storm-Petrel, 2010-2015. Boxplots summarize proportion of recordings with detection of two or more individuals.")} -->
<!-- boxplot(voc.byyear.ftsp.w, 'week', 'prop2', 'species') -->
<!-- ``` -->

