---
title: "Short-term impacts of island restoration on seabirds in Gwaii Haanas"
knit: (function(inputFile, encoding) { 
      out_dir <- 'reports';
      ext <- '.html';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir,  paste('GH-acoustic-Report-', lubridate::date(Sys.time()), ext, sep = ''))) })
output:
  html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo=F, warning=F,message=F,results="hide", prompt=F, error=F, fig.path = 'reports/files/')
```


```{r}
source('header.R')
source('baci-plots.R')
source('baci-prep.R')
source('baci-models.R')
source('baci-captions.R')
```

# BACI

Task 2e) `r task2e`

##### Experimental Design
This study included multiple control and impact sites, across multiple years before and after impact (i.e. rat eradication). ARU monitoring locations ('sites') were located within island groups ('islands'), which were classified as 'control' or 'impact'. The number of sites within islands ranged from 1 to 8. Two response variables were considered: (1) the proportion of recordings with detection of at least one individual across a common within-season time-period (p(presence)); (2) the proportion of recordings with detection of more than one individual (i.e. overlapping calls) across a common within-season time-period (p(presence > 1). Separate BACI analyses were carried out for each species (`r sp[1]`, `r sp[2]`, `r sp[3]`), eradication phase (Phase 1, Phase 2) and response (p(presence), p(presence > 1)).

In phase 1, rat eradication occurred by bait station post-breeding season, 2011. There were two impact islands (Arichika and Bischofs) and three control islands (Alder, Hotspring, Ramsay) and in general, 2 years of sampling... 

In phase 2, eradication of rats took place post-breeding season, 2013. There were two impact sites (Faraday, Murchison) and 4 control sites (Alder, Hotspring/House, Ramsay) and in general, 3 years of sampling before impact (2011, 2012, 2013) and 2 years of sampling after impact (2014, 2015). In total, there were 53 replciate:year combinations (after filtering out replicates that sampled outside of our time-window; see below). Replicates were located at fixed locataions across years. The sampling design was unbalanced; the number of replicates at each site ranged from 1 to 8, and replicates did not consistently sample in all years (2011-2015). 

Variation occurred in sampling duration (i.e. start and end time) of ARUs: across years (e.g. 2011, 2013), within-years (e.g. April 26 - June 06) and within each night (e.g. 12:00am - 3:40am), which made it necessary to constrain our analysis to a common time window, which was justified by observed vocalization probability. We selected the maximum mean vocalization probability within a sliding window of 14 days (within-season) and 1 hour (i.e. 3 recordings; within-night) for each species (Figure ...). Cassin's Auklets had one large peak in vocalization in early April and another (somewhat smaller) in early-mid May; we used a sampling window in early May because many ARUs were not yet operational in early April. We assumed these windows to be representative of the breeding population for that season, i.e. a period of minimal permanent egress/ingress of breeders. Further, since ARUs were often operating in April-June, these windows were generally well represented by monitoring locations in each year.

##### Statistical model
Prior to analysis, each response was logit-transformed (empirical logit transformation) to account for proportions being bounded at 0 and 1. Coefficients, marginal means and BACI contrast and confidence limits were back-transformed after model fitting for presentation in tables. For each species:phase:response combination, we fit a linear mixed-effects model within R, using the lmer() function within the lme4 package. Due to the unbalanced experimental design (e.g. not all monitoring locations consistently monitored in every year), the response was calculated at each monitoring location rather than each island group; this allowed for testing random effects of site, nested within island. For all models, fixed effects included period (Before/After), siteclass (Control/Impact), and period:siteclass interaction; random effects included year, island and site (nested within island). 

We used ANOVA (type 3 with Kenward-Roger approximation for degrees of freedom) to determine statistical significance of fixed effects. The interaction term (period:siteclass) is the 'BACI effect' and indicates a difference in before-after response for control and impact groups. For all coefficients, a p-value < 0.05 indicated statistical significance. We also determined the magnitude of the BACI effect (contrast) by calculating the difference of the differences in marginal means (uCA = uCB - uIA + uIB) along with 95% confidence limits; these analyses were done using the lsmeans package. The contrast can be interpreted as percent increase in the response in the impact area relative the the control area; confidence limits not including 0 indicate an effect.

Key assumptions of the model:    

* System is in equilibirum before and after impact.   

    + This assumption may not be valid as recovery of seabird breeding population may occur over a longer time period than is accounted for in this study.   
    
* Variance is homogeneous, data are normally distributed.   

    + We used several diagnostic plots (residuals vs. fitted values, qqplot for normality) to test validity of the statistical model and associated p-values.    
    
* Observations are independent within each site-year combination and across years.   

    + Sites were spaced sufficeintly far apart in space to warrant assumption of independence.     

## Ancient Murrelet
`r figs("anmu.temp.ph1")`

```{r anmu.temp.ph1)}
tempcov_baci(data = filter(sea, phase1 == 1 & island != 'Hotspring'), x = 'night', y = 'siteID', color = 'exp', date.start = '05-11', date.end = '05-24')
```
  
`r figs("anmu.temp.ph2")`

```{r anmu.temp.ph2)}
tempcov_baci(data = filter(sea, phase2 == 1 & island != 'Hotspring'), x = 'night', y = 'siteID', color = 'exp', date.start = '05-11', date.end = '05-24')
```
  
### Mean proportions by island
`r figs("anmu.mean.ph1")`

```{r, anmu.mean.ph1)}
island_props(data = filter(baci.island, species == sp[1] & phase1 == 1 & island != 'Hotspring'), vline = 2011.5)
```
  
`r figs("anmu.mean.ph2")`

```{r, mean.anmu.ph2)}
island_props(data = filter(baci.island, species == sp[1] & phase2 == 1 & island != 'Hotspring'), vline = 2013.5)
```

### Statistical model results
#### Phase 1, p(presence)
`r tbls("anmu.ph1.1.coef")`

```{r anmu.ph1.1.coef, results = "asis"}
kable(summary_coef(anmu.ph1.1))
```


`r tbls("anmu.ph1.1.marg")`

```{r anmu.ph1.1.marg, results = "asis"} 
knitr::kable(summary_marg(anmu.ph1.1, 'PeriodPh1'))
```
  

`r tbls("anmu.ph1.1.cont")`

```{r anmu.ph1.1.cont, results = "asis"} 
kable(summary_cont(anmu.ph1.1, 'PeriodPh1'))
```
 
 
#### Phase 1, p(presence>1)
`r tbls("anmu.ph1.2.coef")`
```{r anmu.ph1.2.coef, results = "asis"}
kable(summary_coef(anmu.ph1.2))
```

`r tbls("anmu.ph1.2.marg")`
```{r anmu.ph1.2.marg, results = "asis"} 
knitr::kable(summary_marg(anmu.ph1.2, 'PeriodPh1'))
```

`r tbls("anmu.ph1.2.cont")`
```{r anmu.ph1.2.cont, results = "asis"} 
kable(summary_cont(anmu.ph1.2, 'PeriodPh1'))
```

#### Phase 2, p(presence)
`r tbls("anmu.ph2.1.coef")`
```{r anmu.ph2.1.coef, results = "asis"}
kable(summary_coef(anmu.ph2.1))
```

`r tbls("anmu.ph2.1.marg")`
```{r anmu.ph2.1.marg, results = "asis"} 
knitr::kable(summary_marg(anmu.ph2.1, 'PeriodPh2'))
```

`r tbls("anmu.ph2.1.cont")`
```{r anmu.ph2.1.cont, results = "asis"} 
kable(summary_cont(anmu.ph2.1, 'PeriodPh2'))
```

#### Phase 2, p(presence>1)
`r tbls("anmu.ph2.2.coef")`
```{r anmu.ph2.2.coef, results = "asis"}
kable(summary_coef(anmu.ph2.2))
```

`r tbls("anmu.ph2.2.marg")`
```{r anmu.ph2.2.marg, results = "asis"} 
knitr::kable(summary_marg(anmu.ph2.2, 'PeriodPh2'))
```

`r tbls("anmu.ph2.2.cont")`
```{r anmu.ph2.2.cont, results = "asis"} 
kable(summary_cont(anmu.ph2.2, 'PeriodPh2'))
```

## Cassin's Auklet
`r figs("caau.temp.ph1")`

```{r caau.temp.ph1)}
tempcov_baci(data = filter(sea, phase1 == 1), x = 'night', y = 'siteID', color = 'exp', date.start = '05-01', date.end = '05-14')
```
  
`r figs("caau.temp.ph2")`

```{r caau.temp.ph2)}
tempcov_baci(data = filter(sea, phase2 == 1), x = 'night', y = 'siteID', color = 'exp', date.start = '05-01', date.end = '05-14')
```
  
### Mean proportions by island
`r figs("caau.mean.ph1")`

```{r, caau.mean.ph1)}
island_props(data = filter(baci.island, species == sp[2] & phase1 == 1), vline = 2011.5)
```
  
`r figs("caau.mean.ph2")`

```{r, mean.caau.ph2)}
island_props(data = filter(baci.island, species == sp[2] & phase2 == 1), vline = 2013.5)
```

### Statistical model results
#### Phase 1, p(presence)
`r tbls("caau.ph1.1.coef")`
```{r caau.ph1.1.coef, results = "asis"}
kable(summary_coef(caau.ph1.1))
```

`r tbls("caau.ph1.1.marg")`
```{r caau.ph1.1.marg, results = "asis"} 
knitr::kable(summary_marg(caau.ph1.1, 'PeriodPh1'))
```
  
`r tbls("caau.ph1.1.cont")`
```{r caau.ph1.1.cont, results = "asis"} 
kable(summary_cont(caau.ph1.1, 'PeriodPh1'))
```
 
#### Phase 1, p(presence>1)
`r tbls("caau.ph1.2.coef")`
```{r caau.ph1.2.coef, results = "asis"}
kable(summary_coef(caau.ph1.2))
```

`r tbls("caau.ph1.2.marg")`
```{r caau.ph1.2.marg, results = "asis"} 
knitr::kable(summary_marg(caau.ph1.2, 'PeriodPh1'))
```

`r tbls("caau.ph1.2.cont")`
```{r caau.ph1.2.cont, results = "asis"} 
kable(summary_cont(caau.ph1.2, 'PeriodPh1'))
```

#### Phase 2, p(presence)
`r tbls("caau.ph2.1.coef")`
```{r caau.ph2.1.coef, results = "asis"}
kable(summary_coef(caau.ph2.1))
```

`r tbls("caau.ph2.1.marg")`
```{r caau.ph2.1.marg, results = "asis"} 
knitr::kable(summary_marg(caau.ph2.1, 'PeriodPh2'))
```

`r tbls("caau.ph2.1.cont")`
```{r caau.ph2.1.cont, results = "asis"} 
kable(summary_cont(caau.ph2.1, 'PeriodPh2'))
```

#### Phase 2, p(presence>1)
`r tbls("caau.ph2.2.coef")`
```{r caau.ph2.2.coef, results = "asis"}
kable(summary_coef(caau.ph2.2))
```

`r tbls("caau.ph2.2.marg")`
```{r caau.ph2.2.marg, results = "asis"} 
knitr::kable(summary_marg(caau.ph2.2, 'PeriodPh2'))
```

`r tbls("caau.ph2.2.cont")`
```{r caau.ph2.2.cont, results = "asis"} 
kable(summary_cont(caau.ph2.2, 'PeriodPh2'))
```

## Fork-Tailed Storm Petrel
`r figs("ftsp.temp.ph1")`

```{r ftsp.temp.ph1)}
tempcov_baci(data = filter(sea, phase1 == 1), x = 'night', y = 'siteID', color = 'exp', date.start = '05-01', date.end = '05-14')
```
  
`r figs("ftsp.temp.ph2")`

```{r ftsp.temp.ph2)}
tempcov_baci(data = filter(sea, phase2 == 1), x = 'night', y = 'siteID', color = 'exp', date.start = '05-01', date.end = '05-14')
```
  
### Mean proportions by island
`r figs("ftsp.mean.ph1")`

```{r, ftsp.mean.ph1)}
island_props(data = filter(baci.island, species == sp[3] & phase1 == 1), vline = 2011.5)
```
  
`r figs("ftsp.mean.ph2")`

```{r, mean.ftsp.ph2)}
island_props(data = filter(baci.island, species == sp[3] & phase2 == 1), vline = 2013.5)
```

### Statistical model results
#### Phase 1, p(presence)
`r tbls("ftsp.ph1.1.coef")`
```{r ftsp.ph1.1.coef, results = "asis"}
kable(summary_coef(ftsp.ph1.1))
```

`r tbls("ftsp.ph1.1.marg")`
```{r ftsp.ph1.1.marg, results = "asis"} 
knitr::kable(summary_marg(ftsp.ph1.1, 'PeriodPh1'))
```
  
`r tbls("ftsp.ph1.1.cont")`
```{r ftsp.ph1.1.cont, results = "asis"} 
kable(summary_cont(ftsp.ph1.1, 'PeriodPh1'))
```
 
#### Phase 1, p(presence>1)
`r tbls("ftsp.ph1.2.coef")`
```{r ftsp.ph1.2.coef, results = "asis"}
kable(summary_coef(ftsp.ph1.2))
```

`r tbls("ftsp.ph1.2.marg")`
```{r ftsp.ph1.2.marg, results = "asis"} 
knitr::kable(summary_marg(ftsp.ph1.2, 'PeriodPh1'))
```

`r tbls("ftsp.ph1.2.cont")`
```{r ftsp.ph1.2.cont, results = "asis"} 
kable(summary_cont(ftsp.ph1.2, 'PeriodPh1'))
```

#### Phase 2, p(presence)
`r tbls("ftsp.ph2.1.coef")`
```{r ftsp.ph2.1.coef, results = "asis"}
kable(summary_coef(ftsp.ph2.1))
```

`r tbls("ftsp.ph2.1.marg")`
```{r ftsp.ph2.1.marg, results = "asis"} 
knitr::kable(summary_marg(ftsp.ph2.1, 'PeriodPh2'))
```

`r tbls("ftsp.ph2.1.cont")`
```{r ftsp.ph2.1.cont, results = "asis"} 
kable(summary_cont(ftsp.ph2.1, 'PeriodPh2'))
```

#### Phase 2, p(presence>1)
`r tbls("ftsp.ph2.2.coef")`
```{r ftsp.ph2.2.coef, results = "asis"}
kable(summary_coef(ftsp.ph2.2))
```

`r tbls("ftsp.ph2.2.marg")`
```{r ftsp.ph2.2.marg, results = "asis"} 
knitr::kable(summary_marg(ftsp.ph2.2, 'PeriodPh2'))
```

`r tbls("ftsp.ph2.2.cont")`
```{r ftsp.ph2.2.cont, results = "asis"} 
kable(summary_cont(ftsp.ph2.2, 'PeriodPh2'))
```







